Bow shocks, bow waves, and dust waves. IV. Shell shape statistics

### 2019-05-10 papers/Papers1234/app-p-values.tex
New version of abstract for Paper IV

Results from all the statistical tests discussed in
\S~\ref{sec:corr-shape} are given in Table~\ref{tab:big-p}.

% \begin{table}
%   \caption{Lower bounds on the Type~I error rate, \(\alpha\)}
%   \label{tab:type-I}
%   \centering
%   \begin{tabular}{lrr} \toprule
%     Correlation & \(p\) & \(\alpha\) \\
%     \midrule
%     \multicolumn{3}{l}{\itshape Quantitative \dotfill}\\
%     \(R_c\) vs \(R_0\) & \textit{0.0054} & \textit{0.0712} \\
%     \(R_{90}\) vs \(R_0\) & \textbf{0.0001} & \textbf{0.0025} \\
%     \(R_c\) vs \(H_0\) & 0.1229 & 0.4119 \\
%     \(R_{90}\) vs \(H_0\) & \textit{0.0215} & \textit{0.1833} \\
%     \(R_c\) vs \(A_K\) & 0.19 & 0.4617 \\
%     \(R_{90}\) vs \(A_K\) & 0.63 & 1 \\
%     \(R_c\) vs \(|b|\) & 0.19 & 0.4617 \\
%     \(R_{90}\) vs \(|b|\) & 0.31 & 0.4967 \\
%     \(R_c\) vs \(\cos(\ell)\) & 1.00 & 1 \\
%     \(R_{90}\) vs \(\cos(\ell)\) & 0.36 & 0.4999 \\
%     \midrule
%     \multicolumn{3}{l}{\itshape Categorical} \dotfill\\
%     \(R_c\) vs Facing & 0.71 & 1 \\
%     \(R_c\) vs H II & 0.50 & 1 \\
%     \(R_{90}\) vs Facing & 0.60 & 1 \\
%     \(R_{90}\) vs H II & 0.52 & 1 \\
%     \(R_c\) vs Multiple & 1.00 & 1 \\
%     \(R_{90}\) vs Multiple & 0.34 & 0.4993 \\
%     \(R_c\) vs \SI{8}{\um} & 0.82 & 1 \\
%     \(R_{90}\) vs \SI{8}{\um} & 0.6 & 1 \\
%     \midrule
%     \multicolumn{3}{l}{\itshape Other datasets} \dotfill\\
%     \(R_c\) vs Herschel & 0.074 & 0.3437 \\
%     \(R_{90}\) vs Herschel & \textbf{0.00052} & \textbf{0.0106} \\
%     \(R_c\) vs M42 & \textbf{0.00048} & \textbf{0.0099} \\
%     \(R_{90}\) vs M42 & \textbf{0.0000105} & \textbf{0.0003} \\
%     \bottomrule
%   \end{tabular}
% \end{table}

The \(p\)-values are the probability of finding a difference between
two populations as large as (or larger than) what is observed,
\emph{given} that there is no difference in the underlying
distribution from which the two populations are drawn (that is, given
that the null hypothesis is true).  However, what we really want to
know is something else: the probability that the null hypothesis is
true, \emph{given} the observations.  That is, the probability,
\(\alpha\), of a \textit{false positive}, also known as the \textit{Type I
  error rate}.  The common mistake of conflating these two definitions
is known as the ``\(p\)-value fallacy'' \citep{Goodman:1999a}, or``the
error of the transposed conditional'', as discussed in detail by
\citet{Colquhoun:2014a}.  It is possible to derive \(\alpha\) from
\(p\) using Bayes' theorem (e.g., \citealp{Goodman:1999b}), but that
requires an estimate of the prior probability of the null hypothesis,
independent of the observations.  Alternatively, it is also possible
to find a lower bound on \(\alpha\) from a frequentist approach
\citep{Sellke:2001a}:
\begin{equation}
  \label{eq:type-I}
  \alpha(p) \ge \bigg[ 1 - \big(e\, p \ln p\big)^{-1} \bigg]^{-1}
  \quad \text{valid for } p < 1/e.
\end{equation}
This is the approach we adopt here, which also numerically coincides
with the Bayesian approach for the case where the prior probability of
the null hypothesis is 0.5.  The reason that this is only a lower
limit for \(\alpha\) is that if we have overwhelming a priori evidence that
the null hypothesis is true (for instance, from previous empirical
studies, or because it follows from a well-supported theory), then a
Bayesian calculation would give a much higher value of \(\alpha\) than
\eqref{eq:type-I} does.  In our case, however, we have no strong
reasons for favoring any of the null hypotheses, so it is reasonable
to assume \(\alpha\) is close to the lower limit given in \eqref{eq:type-I}.

In order to choose a threshold \(p\)-value that counts as a
``significant'' result, one then needs to balance the risks of false
positives against the risks of \textit{false negatives}.  The false
negative probability, \(\beta\), also known as \textit{Type II error
  rate}, is the probability of failing to reject an untrue null
hypothesis.  That is, in the context of this paper, it is the
probability of failing to detect a real difference between two
sub-samples, or a real correlation between two variables.  The
complementary probability, \(1 - \beta\), is known as the
\textit{statistical power} or sensitivity of the test.  The value of
\(\beta\) depends on three factors:
\begin{enumerate}[1.]
\item The \textit{effect size}, which is a measure of the magnitude of
  the difference in a dependent variable between two sub-samples, or
  the degree of correlation between two continuous variables.  For the
  two sub-sample case, it is common to use a standardised mean
  difference, such as Cohen's \(d\) statistic \citep{Cohen:1988a}:
  \(d = (\bar{X}_A - \bar{X}_B) / s\), where \(\bar{X}_A\),
  \(\bar{X}_B\) are the means of the dependent variable \(X\) for
  samples A and B, while \(s\) is the pooled standard deviation of
  \(X\).  For the case of two continuous variables, the Pearson linear
  correlation coefficient, \(r\), can be used.  In both cases, rules
  of thumb have been developed \citep{Ruscio:2008a} for classifying an
  effect as ``large'' (\(d > 0.8\), \(r > 0.4\)) or ``small''
  (\(d < 0.2\), \(r < 0.1\)).  Alternatively, non-parametric
  statistics can be used, such as the \(A\) measure of stochastic
  superiority \citep{Delaney:2002a}.
\end{enumerate}

Obviously, this depends on the \textit{effect size}, which is the

All astronomical data analysis is \emph{post hoc} analysis, since the universe was not set up to test a particular hypothesis (as far as we know).  It is therefore important to guard against the ``multiple comparisons problem'', whereby seemingly significant correlations are found where none really exist, simply by virtue of the large number of tests that were carried out.

Under the more conservative Holm--Bonferroni method, only comparisons with \(p < 0.001\) would be significant.

The p-curve \citep{Head:2015a}

@@

\section{Perturbed bows}
\label{sec:perturbed-bows}

>>>>

Results from all the statistical tests of the shape distributions
discussed in \S~\ref{sec:corr-shape} are given in
Table~\ref{tab:big-p}.  The \(p\)-values are the probability of
finding a difference between two populations at least as large as what
is observed \emph{given} that there is no difference in the underlying
distribution from which the two populations are drawn (that is, given
that the null hypothesis is true).  Conventionally, the null
hypothesis is rejected at a certain significance threshold \(\alpha\) when
\(p < \alpha\).  Since we are blindly testing many different hypotheses at
once, the commonly used \(\alpha = 0.05\) threshold is too lenient.  In
Figure~\ref{fig:histo-p-values} we analyse the frequency distribution
of \(p\) from all our tests \citep[see][]{Head:2015a}, finding a
systematic excess over a uniform distribution only for \(p < 0.01\).
We therefore take \(\alpha = 0.003\) as the optimum threshold in order to
balance the risks of false positives and false negatives. A false
positive is the erroneous rejection of the null hypothesis (the
spurious detection of a correlation that is not really there), while a
false negative is failing to detect a true correlation.

% However, what we really want to
% know is something else: the probability that the null hypothesis is
% true, \emph{given} the observations.  That is, the probability,
% \(\alpha\), of a \textit{false positive}, also known as the \textit{Type I
%   error rate}.  The common mistake of conflating these two definitions
% is known as the ``\(p\)-value fallacy'' \citep{Goodman:1999a}, or``the
% error of the transposed conditional'', as discussed in detail by
% \citet{Colquhoun:2014a}.  It is possible to derive \(\alpha\) from
% \(p\) using Bayes' theorem (e.g., \citealp{Goodman:1999b}), but that
% requires an estimate of the prior probability of the null hypothesis,
% independent of the observations.  Alternatively, it is also possible
% to find a lower bound on \(\alpha\) from a frequentist approach
% \citep{Sellke:2001a}:
% \begin{equation}
%   \label{eq:type-I}
%   \alpha(p) \ge \bigg[ 1 - \big(e\, p \ln p\big)^{-1} \bigg]^{-1}
%   \quad \text{valid for } p < 1/e.
% \end{equation}
% This is the approach we adopt here, which also numerically coincides
% with the Bayesian approach for the case where the prior probability of
% the null hypothesis is 0.5.  The reason that this is only a lower
% limit for \(\alpha\) is that if we have overwhelming a priori evidence that
% the null hypothesis is true (for instance, from previous empirical
% studies, or because it follows from a well-supported theory), then a
% Bayesian calculation would give a much higher value of \(\alpha\) than
% \eqref{eq:type-I} does.  In our case, however, we have no strong
% reasons for favoring any of the null hypotheses, so it is reasonable
% to assume \(\alpha\) is close to the lower limit given in \eqref{eq:type-I}.

% In order to choose a threshold \(p\)-value that counts as a
% ``significant'' result, one then needs to balance the risks of false
% positives against the risks of \textit{false negatives}.  The false
% negative probability, \(\beta\), also known as \textit{Type II error
%   rate}, is the probability of failing to reject an untrue null
% hypothesis.  That is, in the context of this paper, it is the
% probability of failing to detect a real difference between two
% sub-samples, or a real correlation between two variables.  The
% complementary probability, \(1 - \beta\), is known as the
% \textit{statistical power} or sensitivity of the test.  The value of
% \(\beta\) depends on three factors:
% \begin{enumerate}[1.]
% \item The \textit{effect size}, which is a measure of the magnitude of
%   the difference in a dependent variable between two sub-samples, or
%   the degree of correlation between two continuous variables.  For the
%   two sub-sample case, it is common to use a standardised mean
%   difference, such as Cohen's \(d\) statistic \citep{Cohen:1988a}:
%   \(d = (\bar{X}_A - \bar{X}_B) / s\), where \(\bar{X}_A\),
%   \(\bar{X}_B\) are the means of the dependent variable \(X\) for
%   samples A and B, while \(s\) is the pooled standard deviation of
%   \(X\).  For the case of two continuous variables, the Pearson linear
%   correlation coefficient, \(r\), can be used.  In both cases, rules
%   of thumb have been developed \citep{Ruscio:2008a} for classifying an
%   effect as ``large'' (\(d > 0.8\), \(r > 0.4\)) or ``small''
%   (\(d < 0.2\), \(r < 0.1\)).  Alternatively, non-parametric
%   statistics can be used, such as the \(A\) measure of stochastic
%   superiority \citep{Delaney:2002a}.
% \end{enumerate}

% Obviously, this depends on the \textit{effect size}, which is the

% All astronomical data analysis is \emph{post hoc} analysis, since the universe was not set up to test a particular hypothesis (as far as we know).  It is therefore important to guard against the ``multiple comparisons problem'', whereby seemingly significant correlations are found where none really exist, simply by virtue of the large number of tests that were carried out.

% Under the more conservative Holm--Bonferroni method, only comparisons with \(p < 0.001\) would be significant.

% The p-curve \citep{Head:2015a}

@@

\section{Perturbed bow shocks}
\label{sec:perturbed-bows}

### 2019-05-10 papers/Papers1234/obs-bowshocks.tex
Redo the stats table, and make it portrait orientation
Switch to Kuiper test instead of K-S
New version of abstract for Paper IV

% \title[Stellar bow shocks]{Weather vanes and runaways: Bow
%   shocks across the HR diagram}
\title[Bow shocks, bow waves, and dust waves. IV.]{Bow shocks, bow
  waves, and dust waves. IV. Bow shape statistics}

@@

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Stellar bow shocks are the result of the supersonic interaction
  between a stellar wind and its environment.  Some of these are
  "runaways": high-velocity stars that have been ejected from a star
  cluster.  Others are "weather vanes", where it is the local
  interstellar medium itself that is moving, perhaps as the result of
  a champagne flow of ionized gas from a nearby \hii{} region.
  We recently proposed a new two-dimensional classification scheme for
  the shape of such bow shocks, which we here apply to three very
  different observational datasets: (1)~mid-infrared arcs
  around hot main-sequence stars (\(N = 227\)); (2)~far-infrared arcs
  around luminous cool stars (\(N = 16\)); and (3)~emission-line arcs
  around proplyds and other young stars in the Orion Nebula
  (\(N = 18\)).  We find significant differences between the three
  datasets: cool star bow shocks have markedly more closed wings than
  hot star bow shocks, while differences in the shape of the apex
  region are only marginally significant.  The Orion Nebula arcs, on
  the other hand, have both significantly more open wings and
  significantly flatter apexes than the hot star bow shocks.  We
  discuss the implications of these differences for understanding the
  physics of the bow shock interaction in the three classes of
  sources.
\end{abstract}

@@

% Don't change these lines
\bsp	% typesetting comment
\label{lastpage}

>>>>

% \title[Stellar bow shocks]{Weather vanes and runaways: Bow
%   shocks across the HR diagram}
\title[Bow shocks, bow waves, and dust waves. IV.]{Bow shocks, bow
  waves, and dust waves. IV. Shell shape statistics}

@@

\begin{document}
\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\maketitle
\begin{abstract}
  Stellar bow shocks result from stars that move with respect to their
  environment. The interaction of the stellar wind and radiation with
  gas and dust in the interstellar medium produces curved arcs of
  emission at optical, infrared, and radio wavelengths.  We recently
  proposed a new two-dimensional classification scheme for the shape
  of such bow shocks, which we here apply to three very different
  observational datasets: (1)~mid-infrared arcs around hot OB stars;
  (2)~far-infrared arcs around luminous cool stars; and
  (3)~H\(\alpha\) emission-line arcs around proplyds and other young stars
  in the Orion Nebula.  For OB stars, the average shape is consistent
  with simple thin-shell models for the interaction of a spherical
  wind with a parallel stream, but the diversity of observed shapes is
  many times larger than such models predict.  We propose that this
  may be caused by time-dependent oscillations in the bow shocks, due
  to either instabilities or wind variability.  Cool star bow shocks
  have markedly more closed wings than hot star bow shocks, which may
  be due to the dust emission arising in the shocked stellar wind
  instead of the shocked interstellar medium.  The Orion Nebula arcs,
  on the other hand, have both significantly more open wings and
  significantly flatter apexes than the hot star bow shocks.  We test
  several possible explanations for this difference: divergent ambient
  stream, low Mach number, observational biases, and influence of
  collimated jets.  Further observations are required to decide
  between these.
\end{abstract}

@@

% Don't change these lines

\bsp	% typesetting comment
\label{lastpage}

### 2019-05-10 papers/Papers1234/sec-obs-intro.tex
Finish discussion of OB star shapes for Paper IV
Finish discussion for Paper IV - now just summary to go!
New version of abstract for Paper IV

Stellar bow shocks are produced by the relative motion between a star
and its surrounding medium, and are commonly detected as curved arcs
of emission at optical \citep{Gull:1979a, Brown:2005a}, infrared
\citep{van-Buren:1988a, Kobulnicky:2016a}, or radio
\citep{van-Buren:1990a, Benaglia:2010a} wavelengths.  The canonical
theory for these objects is that they are formed by a two-shock
interaction between the stellar wind and the interstellar medium
\citep{Pikelner:1968a, Dyson:1972a}, which is distorted due to the
supersonic motion of the star \citep{Baranov:1970a, Wilkin:1996a}.  In
some instances, however, the absorbed stellar radiation pressure may
be more important than the stellar wind in providing the inner support
for the bow shell \citep[Paper~I]{Henney:2019a} and this may even be
sufficient to break the collisional coupling between gas and dust
grains \citep[Paper~II]{Henney:2019b}.  In other cases, the appearance
of an infrared emission arc may be due to the illumination of the
inner wall of an asymmetrical cavity \citep{Mackey:2016a}, rather than
the formation of a dense shell, in which case the relative velocity of
the star may be subsonic with respect to its surroundings
\citep{Mackey:2015a}.

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{figs/obs-shape-terminology}
  \caption[Terminology]{Terminology employed in this paper to describe
    bow shock shapes, following \citet{Tarango-Yong:2018a}.}
  \label{fig:obs-shape-terminology}
\end{figure}

@@

In this paper, we investigate the shapes of stellar bow shocks by
calculating the distributions of planitude and alatude for different
classes of bow shock source.  The remainder of the paper is organized
as follows. In \S~\ref{sec:mid-infrared-arcs} we present an analysis
of the shapes of several hundred bowshock candidates associated with
OB stars from the \SI{24}{\um} survey of \citet{Kobulnicky:2016a}.
Our algorithm for automatically fitting and tracing the shapes is
described in \S~\ref{sec:autom-trac-fitt}, together with our ``star
rating'' system for evaluating the fit quality, while in
\S~\ref{sec:ob-shapes} we locate the sources on the planitude--alatude
plane.  In \S~\ref{sec:corr-size} we study the correlations amongst
non-shape parameters of the bow shock sources, such as angular size
and stellar magnitude, while in \S~\ref{sec:corr-shape} we explore the
correlations these parameters and the planitude and alatude.  In
\S~\ref{sec:far-infrared-arcs} we compare with results for bow shocks
around cool luminous stars and in \S~\ref{sec:stat-emiss-line} we
compare with results for staionary emission line arcs in the Orion
Nebula.  In \S~\ref{sec:discussion} we discuss the implications of our
findings for physical models of bow shock formation in the different
classes of sources, and in \S~\ref{sec:conclusion} we summarise our
results.  Further details of the statistical tests that we have
applied are provided in Appendix~\ref{sec:distr-p-values}.

>>>>

Stellar bow shocks are produced by the relative motion between a star
and its surrounding medium, and are commonly detected as curved arcs
of emission at optical \citep{Gull:1979a, Brown:2005a}, infrared
\citep{van-Buren:1988a, Kobulnicky:2016a}, or radio
\citep{van-Buren:1990a, Benaglia:2010a} wavelengths.  The canonical
theory for these objects is that they are formed by a two-shock
interaction between the stellar wind and the interstellar medium
\citep{Pikelner:1968a, Dyson:1972a}, which is distorted due to the
supersonic motion of the star \citep{Baranov:1970a, Wilkin:1996a}.  In
some instances, however, the absorbed stellar radiation pressure may
be more important than the stellar wind in providing the inner support
for the bow shell \citep[Paper~I]{Henney:2019a} and this may even be
sufficient to break the collisional coupling between gas and dust
grains \citep[Paper~II]{Henney:2019b}.  We have proposed a diagnostic
method to distinguish between these cases, based on the bow shock size
and the luminosity ratio between the bow shock and the star
\citep[Paper~III]{Henney:2019c}. In other cases, the appearance of an
infrared emission arc may be due to the illumination of the inner wall
of an asymmetrical cavity \citep{Mackey:2016a}, rather than the
formation of a dense shell, in which case the relative velocity of the
star may be subsonic with respect to its surroundings
\citep{Mackey:2015a}.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.6\linewidth]{figs/obs-shape-terminology}
  \caption[Terminology]{Terminology employed in this paper to describe
    bow shock shapes, following \citet[Paper~0]{Tarango-Yong:2018a}.}
  \label{fig:obs-shape-terminology}
\end{figure}

@@

In this paper, we investigate the shapes of stellar bow shocks by
calculating the distributions of planitude and alatude for different
classes of bow shock source.  The remainder of the paper is organized
as follows. In \S~\ref{sec:mid-infrared-arcs} we present an analysis
of the shapes of several hundred bowshock candidates associated with
OB stars from the \SI{24}{\um} survey of \citet{Kobulnicky:2016a}.
Our algorithm for automatically fitting and tracing the shapes is
described in \S~\ref{sec:autom-trac-fitt}, together with our ``star
rating'' system for evaluating the fit quality, while in
\S~\ref{sec:ob-shapes} we locate the sources on the planitude--alatude
plane.  In \S~\ref{sec:corr-size} we study the correlations amongst
non-shape parameters of the bow shock sources, such as angular size
and stellar magnitude, while in \S~\ref{sec:corr-shape} we explore the
correlations these parameters and the planitude and alatude.  In
\S~\ref{sec:far-infrared-arcs} we compare with results for bow shocks
around cool luminous stars and in \S~\ref{sec:stat-emiss-line} we
compare with results for staionary emission line arcs in the Orion
Nebula.  In \S~\ref{sec:discussion} we discuss the implications of our
findings for physical models of bow shock formation in the different
classes of sources, and in \S~\ref{sec:conclusion} we summarise our
results.  Further details of the statistical tests that we have
applied are provided in Appendix~\ref{sec:distr-p-values} and a simple
model for time-dependent oscillations of the bow shock surface is
presented in Appendix~\ref{sec:perturbed-bows}.
